Typing a linear \picalc

### 2020-03-09 paper/paper.tex
Add process typing rules

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{todonotes}
\usepackage{enumitem}
\setlist{parsep=0pt,listparindent=\parindent}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsbsy}
\usepackage[many]{tcolorbox}
\bibliographystyle{plainurl}

@@

\newcommand{\subst}[3]{#1[#2/#3]}
\newcommand{\picalc}{$\pi$-calculus}
\newcommand{\Picalc}{$\pi$-Calculus}
\newcommand{\reduce}{\longrightarrow}
\newcommand{\types}[4]{#1 \propto #2 \vdash #3 \boxtimes #4}
\newcommand{\suc}{{\scriptscriptstyle 1+}}

% Types
\newcommand{\Var}{\mathrm{VAR}}
\newcommand{\Process}{\mathrm{PROCESS}}
\newcommand{\Unused}{\mathrm{UNUSED}}
\newcommand{\PreCtx}{\mathrm{PRECTX}}
\newcommand{\Ctx}{\mathrm{CTX}}
\newcommand{\Type}{\mathrm{TYPE}\;}
\newcommand{\Idx}{\mathrm{IDX}\;}
\newcommand{\Idxs}{\mathrm{IDXS}}
\newcommand{\Carrier}{\mathrm{CARRIER}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Channel}{\mathrm{CHANNEL}}
\newcommand{\Rec}{\mathrm{REC}}

@@

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    {P \, Q : \Process_n}
    {P \equiv_n Q : Set}}

  \inferrule
  {P \; Q \; R : \Process_n}
  {comp-assoc : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}

  \inferrule
  {P \; Q : \Process_n}
  {comp-sym : \comp{P}{Q} \equiv_n \comp{Q}{P}}

  \inferrule
  {P : \Process_n}
  {comp-end : \comp{P}{\PO_n} \equiv_n P}

  \inferrule
  { }
  {scope-end : \new \PO_{\suc n} \equiv_n \PO_n}

  \inferrule
  {P \, Q : \Process_{\suc n} \\ uQ : \Unused_0 \, Q}
  {scope-ext : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}

  \inferrule
  {P : \Process_{\suc \suc n}}
  {scope-comm : \new \new P \equiv_n \new \new swap \; 0 \; P}
\end{mathpar}
\caption{Structural rewriting rules}
\end{figure}

@@

  \inferrule
  { }
  {zero : \Rec}

  \inferrule
  {r : \Rec}
  {one \; r : \Rec}

  \inferrule
  {r \; s : \Rec}
  {two \; r \; s : \Rec}

@@

  \inferrule
  {P \; Q : \Process_n \\ P \equiv_n Q}
  {struct : P =_{zero} Q}

  \inferrule
  {P \; P' : \Process_{\suc n} \\ P =_r P'}
  {cong-scope : \new P =_{one \; r} \new P'}

  \inferrule
  {P \; P' : \Process_n \\ P =_r P'}
  {cong-comp : \comp{P}{Q} =_{one \; r} \comp{P'}{Q}}

  \inferrule
  {P \; P' : \Process_{\suc n} \\ P =_r P'}
  {cong-recv : \recv{x}P =_{one \; r} \recv{x}P'}

  \inferrule
  {P \; P' : \Process_n \\ P =_r P'}
  {cong-send : \send{x}{y}P =_{one \; r} \send{x}{y}P'}

  \inferrule
  {P : \Process_n}
  {refl : P =_{zero} P}

  \inferrule
  {P \; Q : \Process_n \\ P =_r Q}
  {sym : Q =_{one \; r} P}

  \inferrule
  {P \; Q \; R : \Process_n \\ P =_r Q \\ Q =_s R}
  {trans : P =_{two \; r \; s} R}
\end{mathpar}
\caption{Structural rewriting rules lifted to a congruent equivalence relation}
\end{figure}

@@

  \inferrule
  { }
  {nothing : \Channel_n}

  \inferrule
  {i : \Var_n}
  {just \; i : \Channel_n}

@@

  \inferrule
  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
  {comm : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}

  \inferrule
  {P \reduce_i P'}
  {par : \comp{P}{Q} \reduce_i \comp{P'}{Q}}

  \inferrule
  {P \reduce_i Q}
  {res : \new P \reduce_{dec\; i} \new Q}

  \inferrule
  {P = P' \\ P' \reduce_i Q}
  {struct : P \reduce_i Q}
\end{mathpar}
\caption{Operational semantics indexed by reducing channel}
\end{figure}

@@

@@

  \inferrule
  {\imath : \Idxs_n \\ x : \Idx}
  {\imath , x : \Idxs_{\suc n}}

@@

  {\mprset{fraction={===}}
    \inferrule
    {\imath : \Idxs n}
    {\Ctx_\imath : Set}}

@@

  \inferrule
  {\Gamma : \Ctx_\imath \\ c : \Carrier_x}
  {\Gamma , c : \Ctx_{\imath , x}}
\end{mathpar}
\end{figure}
two-layered approach: types on one hand, capabilities on the other
removing from context vs keeping in context but marking it used

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule{
      \gamma : \PreCtx_n \\
      \inferrule*{}{
        \imath : \Idxs_n \\\\
        \Gamma : \Ctx_\imath} \\
      P : \Process_n \\
      \Delta : \Ctx_\imath}
    {\types{\gamma}{\Gamma}{P}{\Delta} : Set}}

\end{mathpar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject reduction}

@@

Suffices to model as a series of insertions.

>>>>

\usepackage[utf8]{inputenc}
\usepackage{url}
\usepackage{todonotes}
\usepackage{enumitem}
\setlist{parsep=0pt,listparindent=\parindent}
\usepackage{amssymb}
\usepackage[fleqn]{amsmath}
\usepackage{amsbsy}
\usepackage[many]{tcolorbox}
\bibliographystyle{plainurl}

@@

\newcommand{\subst}[3]{#1[#2/#3]}
\newcommand{\picalc}{$\pi$-calculus}
\newcommand{\Picalc}{$\pi$-Calculus}
\newcommand{\reduce}{\longrightarrow}
\newcommand{\types}[4]{#1 \propto #2 \vdash #3 \boxtimes #4}
\newcommand{\contains}[5]{#1 \propto #2 \ni #3 \propto #4 \boxtimes #5}
\newcommand{\suc}{{\scriptscriptstyle 1+}}
\newcommand{\op}[3]{#1 := #2 \cdot #3}

% Types
\newcommand{\Var}{\mathrm{VAR}}
\newcommand{\Process}{\mathrm{PROCESS}}
\newcommand{\Unused}{\mathrm{UNUSED}}
\newcommand{\PreCtx}{\mathrm{PRECTX}}
\newcommand{\Ctx}{\mathrm{CTX}}
\newcommand{\Type}{\mathrm{TYPE}\;}
\newcommand{\Idx}{\mathrm{IDX}\;}
\newcommand{\Idxs}{\mathrm{IDXS}}
\newcommand{\Carrier}{\mathrm{CARRIER}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Channel}{\mathrm{CHANNEL}}
\newcommand{\Rec}{\mathrm{REC}}
\newcommand{\Quantifier}{\mathrm{QUANTIFIER}}
\newcommand{\Quantifiers}{\mathrm{QUANTIFIERS}}
\newcommand{\constr}[1]{\mathtt{#1}}

@@

\subsection{Notation}

\begin{figure}[h]
  \begin{mathpar}
    {\mprset{fraction={===}}
      \inferrule
      { }
      {\N : Set}}

    \inferrule
    { }
    {0 : \N}

    \inferrule
    {n : \N}
    {\suc n : \N}
  \end{mathpar}
  \caption{Notation used in this paper}
\end{figure}

Double rule for type-level definitions.
We omit universe levels for brevity.
Constructors are $\constr{teletyped}$ unless they are symbols.
Some constructors are shared across types. They can however always be disambiguated through the type of the goal.

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule
    { }
    {P \equiv_n Q : Set}}

  \inferrule
  { }
  {\constr{comp-assoc} : \comp{P}{\comp{Q}{R}} \equiv_n \comp{\comp{P}{Q}}{R}}

  \inferrule
  { }
  {\constr{comp-sym} : \comp{P}{Q} \equiv_n \comp{Q}{P}}

  \inferrule
  { }
  {\constr{comp-end} : \comp{P}{\PO_n} \equiv_n P}

  \inferrule
  { }
  {\constr{scope-end} : \new \PO_{\suc n} \equiv_n \PO_n}

  \inferrule
  {P \, Q : \Process_{\suc n} \\ uQ : \Unused_0 \, Q}
  {\constr{scope-ext} : \new (\comp{P}{Q}) \equiv_n \comp{(\new P)}{lower \; 0 \; \; Q \; uQ}}

  \inferrule
  {P : \Process_{\suc \suc n}}
  {\constr{scope-comm} : \new \new P \equiv_n \new \new swap \; 0 \; P}
\end{mathpar}
\caption{Structural rewriting rules indexed by recursion tree. $P \; Q \; R : \Process_n$ where $P$, $Q$, or $R$ have been omitted for brevity.}
\end{figure}

@@

  \inferrule
  { }
  {\constr{zero} : \Rec}

  \inferrule
  {r : \Rec}
  {\constr{one} \; r : \Rec}

  \inferrule
  {r \; s : \Rec}
  {\constr{two} \; r \; s : \Rec}

@@

  \inferrule
  {P \equiv_n Q}
  {\constr{struct} : P =_{\constr{zero}} Q}

  \inferrule
  {P \; P' : \Process_{\suc n} \\ P =_r P'}
  {\constr{cong-scope} : \new P =_{\constr{one} \; r} \new P'}

  \inferrule
  {P =_r P'}
  {\constr{cong-comp} : \comp{P}{Q} =_{\constr{one} \; r} \comp{P'}{Q}}

  \inferrule
  {P \; P' : \Process_{\suc n} \\ P =_r P'}
  {\constr{cong-recv} : \recv{x}P =_{\constr{one} \; r} \recv{x}P'}

  \inferrule
  {P =_r P'}
  {\constr{cong-send} : \send{x}{y}P =_{\constr{one} \; r} \send{x}{y}P'}

  \inferrule
  {P : \Process_n}
  {\constr{refl} : P =_{\constr{zero}} P}

  \inferrule
  {P =_r Q}
  {\constr{sym} : Q =_{\constr{one} \; r} P}

  \inferrule
  {P =_r Q \\ Q =_s R}
  {\constr{trans} : P =_{\constr{two} \; r \; s} R}
\end{mathpar}
\caption{Structural rewriting rules lifted to a congruent equivalence relation. $P \; P' \; Q\; R : \Process_n$ where $P$, $P'$, $Q$, or $R$ have been omitted for brevity.}
\end{figure}

@@

  \inferrule
  { }
  {\constr{nothing} : \Channel_n}

  \inferrule
  {i : \Var_n}
  {\constr{just} \; i : \Channel_n}

@@

  \inferrule
  {i \; j : \Var_n \\ P : \Process_{\suc n} \\ Q : \Process_n \\ uP : \Unused_0 \, P}
  {\constr{comm} : \comp{\recv{i}P}{\send{i}{j}{Q}} \reduce_{just \; i} \comp{lower \; 0 \; \subst{P}{j}{0} \; uP}{Q}}

  \inferrule
  {P \reduce_i P'}
  {\constr{par} : \comp{P}{Q} \reduce_i \comp{P'}{Q}}

  \inferrule
  {P \reduce_i Q}
  {\constr{res} : \new P \reduce_{dec\; i} \new Q}

  \inferrule
  {P = P' \\ P' \reduce_i Q}
  {\constr{struct} : P \reduce_i Q}
\end{mathpar}
\caption{Operational semantics indexed by reducing channel}
\end{figure}

@@

\begin{figure}[h]
\begin{equation}
\begin{split}
  \constr{0\cdot}          &: C \\
  \constr{+\cdot}          &: C \\
  \constr{-\cdot}          &: C \\
  \constr{\op{\_}{\_}{\_}} &: C \to C \to C \to Set \\
  \constr{1\cdot}          &: C \\
  \constr{\cdot-join}      &: \forall \{x y z\} \to \op{x}{y}{+\cdot} \to \op{x}{z}{-\cdot} \to \exists w . (\op{x}{w}{1\cdot}) \\
  \constr{\cdot-compute}   &: \forall y z \to Dec (\exists x . (\op{x}{y}{z})) \\
  \constr{\cdot-compute^l} &: \forall x z \to Dec (\exists y . (\op{x}{y}{z})) \\
  \constr{\cdot-unique}    &: \forall \{x x' y z\} \to \op{x'}{y}{z} \to \op{x}{y}{z} \to x' \equiv x \\
  \constr{\cdot-unique^l}  &: \forall \{x y y' z\} \to \op{x}{y'}{z} \to \op{x}{y}{z} \to y' \equiv y \\
  \constr{\cdot-id^l}      &: \forall x \to \op{x}{0\cdot}{x} \\
  \constr{\cdot-comm}      &: \forall \{x y z\} \to \op{x}{y}{z} \to \op{x}{z}{y} \\
  \constr{\cdot-assoc}     &: \forall \{x y z u v\} \to \op{x}{y}{z} \to \op{y}{u}{v} \to \exists w . (\op{x}{u}{w} \times \op{w}{v}{z})
\end{split}
\end{equation}
\caption{Partial commutative monoid}
\end{figure}
\todo{alignment, get rid of $1\cdot$, $\cdot-join$, $\cdot-compute^l$}

\begin{figure}[h]
\begin{equation}
\begin{split}
  \constr{\Idx}          &: Set \\
  \constr{\exists \Idx}  &: Idx \\
  \constr{\Carrier}      &: \Idx \to Set \\
  \constr{\Quantifiers}  &: \forall i : \Idx \to \Quantifier_{\Carrier_i}
\end{split}
\end{equation}
\caption{Indexed set of partial commutative monoids}
\end{figure}

\subsubsection{Example type systems}

Shared. Gradual. Affine. Linear.

@@

  \inferrule
  {is : \Idxs_n \\ i : \Idx}
  {is , i : \Idxs_{\suc n}}

@@

  {\mprset{fraction={===}}
    \inferrule
    {is : \Idxs_n}
    {\Ctx_{is} : Set}}

@@

  \inferrule
  {\Gamma : \Ctx_{is} \\ x : \Carrier_i}
  {\Gamma , x : \Ctx_{is , i}}
\end{mathpar}
\end{figure}
two-layered approach: types on one hand, capabilities on the other
removing from context vs keeping in context but marking it used

@@

\begin{figure}[h]
\begin{mathpar}
  {\mprset{fraction={===}}
    \inferrule{
      \gamma : \PreCtx_n \\
      \inferrule*{}{
        is : \Idxs_n \\\\
        \Gamma : \Ctx_{is}} \\
      t : \Type \\
      x : \Carrier_i \\
      \Delta : \Ctx_{is}}
    {\contains{\gamma}{\Gamma}{t}{x}{\Delta} : Set}}

  \inferrule
  {\Gamma : \Ctx_{is} \\ y \; z : \Carrier_i \\ True (\constr{\cdot-compute} \; y \; z)}
  {\constr{zero} : \contains{\gamma , t}{\Gamma , x}{t}{y}{\Gamma , z}}

  \inferrule
  {\Gamma : \Ctx_{is} \\ x : \Carrier_i \\ x' : \Carrier_j \\ \Delta : \Ctx_{is} \\\\
    loc_x : \contains{\gamma}{\Gamma}{t}{x}{\Delta}}
  {\constr{suc} : \contains{\gamma , t}{\Gamma , x'}{t}{x}{\Delta , x'}}
\end{mathpar}
\end{figure}

  {\mprset{fraction={===}}
    \inferrule{
      \gamma : \PreCtx_n \\
      \inferrule*{}{
        is : \Idxs_n \\\\
        \Gamma : \Ctx_{is}} \\
      P : \Process_n \\
      \Delta : \Ctx_{is}}
    {\types{\gamma}{\Gamma}{P}{\Delta} : Set}}

  \inferrule
  { }
  {\constr{end} : \types{\gamma}{\Gamma}{\PO}{\Gamma}}

  \inferrule
  {t : \Type \\ x : \Carrier_i \\ y : \Carrier_j \\\\
   cont : \types{\gamma , C[ t \propto x ]}{\Gamma , y}{P}{\Delta , 0.}}
  {\constr{chan} : \types{\gamma}{\Gamma}{\new P}{\Delta}}

  \inferrule
  {chan_x : \contains{\gamma}{\Gamma}{C[ t \propto x]}{+\cdot}{\Xi} \\\\
   cont : \types{\gamma , t}{\Xi , x}{P}{\Theta , 0\cdot}}
  {\constr{recv} : \types{\gamma}{\Gamma}{\recv{toFin \; chan_x}{P}}{\Theta}}

  \inferrule
  {chan_x : \contains{\gamma}{\Gamma}{C[t \propto x]}{-\cdot}{\Delta} \\\\
   loc_y : \contains{\gamma}{\Delta}{t}{x}{\Xi} \\\\
   cont : \types{\gamma}{\Xi}{P}{\Theta}}
  {\constr{send} : \types{\gamma}{\Gamma}{\send{toFin \; chan_x}{toFin \; loc_y}P}{\Theta}}

  \inferrule
  {left : \types{\gamma}{\Gamma}{P}{\Delta} \\\\
    right : \types{\gamma}{\Delta}{Q}{\Xi}}
  {\constr{comp} : \types{\gamma}{\Gamma}{\comp{P}{Q}}{\Xi}}
\end{mathpar}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Subject reduction}

@@

Order preserving embeddings model a series of insertions. We only ever need one insertion to prove subject congruence, but there is no loss of generality.
